---
pagetitle: "sf"
editor_options: 
  chunk_output_type: console
---

# Using the `sf` Package

This package is great and I've already shown a good bit on how to read files in, plot, crop data, etc.

This is meant to link to the existing vignettes, and walk through a few more of the features/functions that `sf` provides. It's really pretty amazing.

## Existing vignettes:

There are some great [vignettes available for `sf`](https://cran.r-project.org/web/packages/sf/index.html). I'd start by looking though [**1**](https://cran.r-project.org/web/packages/sf/vignettes/sf1.html) & [**2**](https://cran.r-project.org/web/packages/sf/vignettes/sf2.html).

### **`sf`** Functions

I think one thing to keep in mind is the `sf` objects are dataframes...but some functions will return vectors or matrices based on the data. Here's a quick comparison of which returns what:

**RETURNS NEW GEOMETRY (so `data.frame`)**
 
 - `st_buffer`
 - `st_boundary` 
 - `st_convexhull`
 - `st_union_cascaded` 
 - `st_simplify` 
 - `st_triangulate`
 - `st_polygonize`
 - `st_centroid`
 - `st_segmentize`
 - `st_union`

**RETURNS index or logical Matrix**

 - `st_intersects`
 - `st_disjoint`
 - `st_touches`
 - `st_crosses`
 - `st_within` 
 - `st_contains`
 - `st_overlaps`
 - `st_equals`
 - `st_covers`
 - `st_covered_by`
 - `st_equals_exact`
 - `st_is_within_distance`
 
### Load Point Data

Let's use our CDEC cnow course data to start, this should return a dataframe with 261 points, each with some basic lat/lon information, etc.

```{r getcdec, eval=T, echo=T}
library(sf)
library(sharpshootR)  # CDEC.snow.courses, CDECquery, CDECsnowQuery
library(dplyr)

# GET DATA
data(CDEC.snow.courses)

# Make into a dataframe
snw<-CDEC.snow.courses
snw$course_number <- as.factor(snw$course_number)
snw$longitude<-as.numeric(snw$longitude)*-1 # fix longitude, must be negative for northern hemisphere
# rename col:
snw <- snw %>% rename(apr1avg_in=april.1.Avg.inches)

# summary of data:
summary(snw)
dim(snw)

```

How do we make the data spatial with `sf`? See example in XY or in Lines/Polygons.

```{r, eval=T, echo=F}
snw_sf <- st_as_sf(snw, 
                   coords = c(6, 5), # can use numbers here too
                   remove = F, # don't remove these lat/lon cols from df
                   crs = 4326) # add projection (this is UTM11)
```

Let's do some basic data viewing/summary.

```{r, eval=T, echo=T}

# get geometry info: returns list column
snw_geom <- st_geometry(snw_sf)

# get coordinates: note this will return a matrix/vector of coordinates
head(snw_coords <- st_coordinates(snw_sf))

st_crs(snw_sf) # check the crs/projection info

```

### Load Shapefile Data: Unzipped

Ok, let's load a shapfile. I'll show how to do it as the unzipped version (standard), and then how to do it with a zipped shapefile.

```{r shapefilepolys}

# notice the similar function structure, but results in dataframe
rivers_sf <- st_read(dsn = "./data", layer="rivs_CA_OR_hydroshed") 

# check crs
st_crs(rivers_sf)

plot(st_geometry(rivers_sf), col="blue3")

```

### Load Shapefile Data: Unzipped

Ok, let's load a shapfile. I'll show how to do it as the unzipped version (standard), and then how to do it with a zipped shapefile.

```{r shapefilepolysZip}

# link to zipped version is here:
# download.file(url = "https://github.com/ryanpeek/mapping_in_R/blob/master/data/rivs_CA_OR_hydroshed.zip?raw=true",
#              destfile = "data/rivs_CA_OR_hydroshed.zip")

# will give a warning message...that's ok
rivers_sf <- st_read(unzip("data/rivs_CA_OR_hydroshed.zip")) 

# now remove the "temp" unzipped shapefile from your directory, 
# Make sure you are removing the temporary unzipped version, and not the original zip.
# the zipped version should remain in "data", the temp unzipped version should be in your main working dir

file.remove(list.files(pattern="rivs_CA_OR_hydroshed", recursive = F)) # should return TRUE for every file associated with shapefile

# check crs
st_crs(rivers_sf)

```

Sweet. Ok...let's show a few other helpful pieces.

## Spatial Operations

Often you need to buffer out from a set of polygons or points...or clip/crop data to a specific polygon. Let's see how that works. Let's use the `USAboundaries` package. Another great package [`tidycensus`](https://walkerke.github.io/tidycensus/), though you need to sign up for a free API key. .


### `USAboundaries` Example: Add Centroid

First get the data and add a centroid to the counties.

```{r usaboundaries, eval=T, echo=T}

library(USAboundaries)

# Pick a State
state_names <- c("california")

# Download STATE data and add projection
CA<-us_states(resolution = "high", states = state_names) %>%
  st_transform(crs = 4269) # defaults to 4326

st_crs(CA)

library(purrr)

# Pick some CA counties
co_names <- c("Butte", "Placer", "El Dorado", "Nevada", "Yuba", "Sierra", "Plumas")

# get COUNTY data for a given state
counties_spec <- us_counties(resolution = "low", states=state_names) %>% # use list of state(s) here
  st_transform(crs=26911) %>% 
  filter(name %in% co_names) # filter to just the counties we want

# to add centroid points to existing sf dataframe, we can use purrr:
counties_spec <- counties_spec %>% 
  mutate(lon=map_dbl(geometry, ~st_centroid(.x)[[1]]), # add centroid values for labels
         lat=map_dbl(geometry, ~st_centroid(.x)[[2]])) # add centroid values for labels

# or to create new sf object, just call st_centroid on dataframe:
counties_centroid <- st_centroid(counties_spec)

plot(st_geometry(counties_spec), border="darkgreen")

# option 1:
plot(st_geometry(counties_centroid), col="maroon", pch=16, add=T)

# option 2:
points(counties_spec$lon, counties_spec$lat, col="blue", pch=21, cex=2)

```

### `tidycensus` Example Data

```{r tidycensus, eval=T, echo=F}

library(tidycensus)
#m90 <- get_decennial(geography = "state", variables = "H043A001", year = 1990)

library(tidyverse)
options(tigris_use_cache = TRUE)

# Pick some CA counties
co_names <- c("Butte", "Placer", "El Dorado", "Nevada", "Yuba", "Sierra", "Plumas")

counties_cens <- get_acs(state = "CA", county = co_names, geography = "tract", 
                  variables = "B19013_001", geometry = TRUE)

#head(counties_cens)

library(viridis)

counties_cens %>%
  ggplot(aes(fill = estimate, color = estimate)) + 
  geom_sf() + 
  coord_sf(crs = 26911) + # NAD83 Zone 11 UTMs metric
  scale_fill_viridis("Median Household Income", option = "magma") + 
  scale_color_viridis("Median Household Income", option = "magma") +
  theme_bw(base_family = "Roboto Condensed")

```

### Crop Data

Let's crop our river data to the counties of interest. Remember, these are dataframes, so we can crop all, or select a single county to crop to. For these operations, it's important things are in same projection.

```{r crop_all, eval=T, echo=T}

# check crs:
st_crs(rivers_sf)
st_crs(counties_spec)

# fix the rivers_sf layer to match counties
rivers_sf <- st_transform(rivers_sf, crs = 26911)

# now crop!
rivers_crop <- st_intersection(rivers_sf, counties_spec)

# plot it
plot(st_geometry(rivers_crop), col="blue") # a quick plot check to make sure it worked

```

What if we want to crop by a single county?

```{r crop_single, eval=T, echo=T}

# check crs:
st_crs(rivers_sf)
st_crs(counties_spec)

# fix the rivers_sf layer to match counties
rivers_sf <- st_transform(rivers_sf, crs = 26911)

# now crop!
rivers_crop1 <- st_intersection(rivers_sf, counties_spec[1,]) # could use counties_spec[counties_spec$Nevada,] too

plot(st_geometry(counties_spec), lwd=2)
plot(st_geometry(rivers_crop), col="blue", add=T) 
plot(st_geometry(rivers_crop1), col="maroon", lwd=2, add=T)

```

### Buffer Data

```{r buffer, echo=T, eval=T}

county_buff <- st_buffer(counties_spec[counties_spec$name=="Nevada",], dist =3000) # in meters

# plot
plot(st_geometry(counties_spec), axes=T)
plot(st_geometry(county_buff), border="red", lwd=2, add=T)
plot(st_geometry(counties_centroid[1,]), pch=16, add=T, col="blue", cex=3)
```

```{r addsnowpoints}

# transform data of snowstations:
st_crs(snw_sf)
snw_sf <- st_transform(snw_sf, crs = 26911)

plot(st_geometry(counties_spec[1,]))
plot(st_geometry(county_buff), border="red", lwd=2, add=T)
plot(st_geometry(counties_centroid[1,]), pch=16, add=T, col="blue", cex=3)
plot(st_geometry(snw_sf), bg="yellow", pch=21, add=T)

```

### Dissolve Boundaries

One other common task is dissolving boundaries across a common group. For example, let's say we want to dissolve al the selected counties into one big polygon.

```{r union, echo=T, eval=T}

counties_dissolve <- st_union(counties_spec,by_feature = FALSE)

# by feature allows all geometries to be unioned together into a single geometry. 
plot(st_geometry(counties_dissolve))

```

## `ggmap`

This allows us to add some boundaries to a map. There are many types of backgrounds. In order for `sf` objects to work within `ggmap`, one small trick you'll need to add is adding `inherit.aes=FALSE` inside the `geom_sf()` call.

```{r eval=T, echo=T}

library(ggmap) # need this package

# ggmap is in WGS84, so need to transform data to get the correct centroid point in lat/lon
counties_dissolve <- counties_dissolve %>% st_transform(crs=4326)
loc_cent <- counties_dissolve  %>% st_centroid() %>% st_coordinates() 

# set mapRange to the dissolved counties for mapping xlim and ylim
mapRange1 <- c(range(st_coordinates(counties_dissolve)[,1]),
               range(st_coordinates(counties_dissolve)[,2]))

# set the center of the map
location=c(loc_cent[1], loc_cent[2]) 

# set the background map: black and white terrain
map1 <- get_map(location=location, crop = F,
             color="bw",
             maptype="terrain",
             source="google",
             zoom=8)

# stamen toner
map2 <- get_map(location=location, crop = F,
             maptype="toner",
             source="stamen",
             zoom=8)

# water color
map3 <- get_map(location=location, crop = F, 
             maptype="watercolor",
             source="stamen",
             zoom=7)

# satellite
map4 <- get_map(location=location,crop = F,
             maptype="hybrid",
             source="google",
             zoom=7)

# make a multiplot with cowplot
library(cowplot)
plot_grid(ggmap(map1), ggmap(map2), 
          ggmap(map3), ggmap(map4), 
          labels = c("Terrain","Toner","Watercolor","Satellite"),
          ncol = 2, align = 'v',label_size = 10, label_fontfamily = "Roboto Condensed")


# now let's make a ggmap with our data
sitemap <- ggmap(map1) # start a ggmap... like calling ggplot()

# and let's make our nicemap...remember to use "inherit.aes=FALSE" for the geom_sf layers
nicemap<-
  sitemap + 
  geom_sf(data=rivers_crop, col="blue", alpha=0.8, size=0.5, inherit.aes = FALSE)+
  geom_sf(data=counties_spec, fill = NA, show.legend = F, color="gray50", lwd=0.4, inherit.aes = FALSE) +
  geom_sf(data=snw_sf, aes(fill=apr1avg_in), pch=21, alpha=0.9, size=3.5, inherit.aes = FALSE) +
  coord_sf(crs = 4326, xlim = mapRange1[c(1:2)], ylim = mapRange1[c(3:4)]) +
  scale_fill_viridis_c(option = "A")+
  labs(x="Longitude (WGS84)", y="Latitude",
       title="Snow Stations of CA") + 
  theme_bw(base_family = "Roboto Condensed") # change this to sans if it doesn't plot
nicemap

```

## `mapview`

A quick and easy way to view our `sf` data in a leaflet map uses a package called `mapview`. Great thing is it's very simple and quick to make an interactive map.

```{r eval=T, echo=T}
library(mapview)

mapview(snw_sf)
```

Add additional layers by calling different maps and adding together in ggplot style:

```{r, eval=T, echo=T}

m1 <- mapview(snw_sf, zcol="apr1avg_in", legend=TRUE)
m2 <- mapview(rivers_crop)
m3 <- mapview(counties_spec, fill=NA, color="maroon")

m1 + m2 + m3

```

# Summary

Hope this helps at least introduce some of the things you can do with `sf` and maps. There are many more functions available to use, so definitely check out the vignettes and resources.
