---
title: ""
---

<br>

# Using `mapview`

This package is pretty cool, and has a bunch of functionality. There are a number of [vignettes online already](https://r-spatial.github.io/mapview/articles/index.html), in particular check out the basic overview vignette [here](https://r-spatial.github.io/mapview/articles/articles/mapview_01-basics.html). 

This tutorial is just a quick demo of what you can do with `mapview`, and to show it's simplicity. It's not meant to be comprehensive. See the vignettes details.

```{r}
library(mapview)
mapview(breweries)
```

<br>

## Main Functions

`mapview` - view potentially multiple spatial objects on a set of background maps
`viewExtent` - view extent / bounding box of spatial objects
`viewRGB` - view RGB true- or false-color images of raster objects
`latticeView` & `sync` - view small multiples of several potentially synchronised maps
`slideView` - view two overlaid raster objects with a horizontal slider to hide/reveal one of them
`cubeView` - 3D view of raster data cubes (e.g. space-time cube) with keyboard control over position of slices
`plainview` - view large rasters on a plain background (i.e. no background map)
`mapshot` - easily save maps (including leaflet maps) as html or png (or other image formats)

### Spatial Data Supported in `mapview`

 - `sf` (XY dimension only)
 - `raster` (Layer, Stack, Brick and SpatialPixels* / SpatialGridDataFrame)
 - `sp` (Points, Polygons, Lines and their DataFrame version)
 - `satellite`

### Leaflet Compatible Functions

In addition, mapview provides some extra leaflet compatible functionality:

 - `addHomeButton` - add zoom-to-layer buttons
 - `addLogo` - add a logo (or any image really) to a map
 - `addFeatures` - add features to a map (sf or sp). This is features type agnostic, meaning the same function can be used for points, lines and polygons
 - `addMouseCoordinates` - add mouse cursor postion information when moving over the map
 - `addLargeFeatures` - add large data to a map that would otherwise make the leaflet map irresponsive
 - `popupTable, popupGraph & popupImage` - popups of tables, grpahs and images for use with the leaflet add* functions

## Fancy `mapview`

https://r-spatial.github.io/mapview/articles/articles/mapview_05-extras.html

A few of the "*fancier*"" things you can do.

<br>

### `slideView`

```{r slideview1}

library(raster)
library(mapview)

stck1 <- subset(poppendorf, c(3, 4, 5))
stck2 <- subset(poppendorf, c(2, 3, 4))
slideView(stck1, stck2)

```

<br>
<br>

```{r slideview2, echo=T, eval=T}

library(jpeg)
library(raster)

web_img2000 <- "http://cdn.newsapi.com.au/image/v1/68565a36c0fccb1bc43c09d96e8fb029"

jpg2000 <- readJPEG(readBin(web_img2000, "raw", 1e6))

# Convert imagedata to raster
rst_blue2000 <- raster(jpg2000[, , 1])
rst_green2000 <- raster(jpg2000[, , 2])
rst_red2000 <- raster(jpg2000[, , 3])

img2000 <- brick(rst_red2000, rst_green2000, rst_blue2000)

web_img2013 <- "http://cdn.newsapi.com.au/image/v1/5707499d769db4b8ec76e8df61933f2a"

jpg2013 <- readJPEG(readBin(web_img2013, "raw", 1e6))

# Convert imagedata to raster
rst_blue2013 <- raster(jpg2013[, , 1])
rst_green2013 <- raster(jpg2013[, , 2])
rst_red2013 <- raster(jpg2013[, , 3])

img2013 <- brick(rst_red2013, rst_green2013, rst_blue2013)

slideView(img2000, img2013, label1 = "before", label2 = "after")

```

Whoa! That's cool.

### `cubeView`

```{r echo=T, eval=F}
library(raster)
library(mapview)

kili_data <- system.file("extdata", "kiliNDVI.tif", package = "mapview")
kiliNDVI <- stack(kili_data)

cubeView(kiliNDVI)

library(RColorBrewer)
clr <- colorRampPalette(brewer.pal(9, "BrBG"))
cubeView(kiliNDVI, at = seq(-0.15, 0.95, 0.1), col.regions = clr)

```

### `sync` views

```{r, echo=T, eval=T}

m1 <- mapview(breweries, zcol = "village", map.types = "Esri.WorldImagery")
m2 <- mapview(breweries, zcol = "brewery", col.regions = heat.colors)
m3 <- mapview(breweries, zcol = "founded", legend = TRUE)
m4 <- leaflet() %>% addTiles() %>% addCircleMarkers(data = breweries)

sync(m1, m2, m3, m4)

sync(list(m1, m2, m3, m4),
     sync = list(c(1, 4),
                 c(2, 3)))
```



