---
title: ""
---

<br>

# Line and Polygon Data with `rgdal` and `sf`

There are loads of spatial mapping/plotting packages in R. The main two ways to read in spatial data use the `rgdal` package, and the `sf` package. Let's look at how to load/plot line and polygon data.

**Let's load packages first:**

```{r load libs, warning=FALSE}
suppressPackageStartupMessages({
  library(rgdal); # spatial/shp reading
  library(dplyr); # wrangling data/plotting
  library(readr);
  library(viridis); # nice color palette
  library(sf); # newer "simple features" spatial package
  library(USAboundaries); # state/county data
  library(Imap); # nice mapping/color functions
  library(ggrepel) # for labeling
  }) 

```

## Line/Polyline Shapefile Data

For this example, let's use some [Hydroshed](http://www.hydrosheds.org/) Data.

I've downloaded rivers for CA and OR and put them [here](https://github.com/ryanpeek/mapping_in_R/blob/master/data/rivs_CA_OR_hydroshed.zip) on github. Download the zipped file and unzip it in a `data` folder. We're going to use *`shapefiles`* for the remainder of this example. 

### Load shapefiles with `rgdal`

Let's load a polyline or line shapefile of rivers of California and Oregon. The result is a `SpatialLinesDataFrame` in your R environment.

```{r shapesRGDAL, echo=T, eval=T}

# we can use ogrInfo to see CRS, attributes, etc.
ogrInfo(dsn="./data", layer="rivs_CA_OR_hydroshed") # see shapefile info

# then read in the shapefile
rivers_sp<- readOGR(dsn = "data", layer = "rivs_CA_OR_hydroshed")

proj4string(rivers_sp) # check projection, should be WGS84

# raster time comparison (uses rgdal currently)
# library(raster)
# rivers_rast <- shapefile("data/rivs_CA_OR_hydroshed.shp")

```

### Load shapefiles with `sf`

Here's how to do the same thing using the `sf` package. Notice two important differences, the `sf` package reads things in as a regular dataframe, with the spatial component of the data contained inside a `geometry` **list-column** within the dataframe. That means you can operate on this data as you would any data frame. The other main difference, is that reading shape data in is much faster with `sf`.  

```{r shapes_sf, echo=T, eval=T}

# notice the similar function structure, but results in dataframe
rivers_sf <- st_read(dsn = "./data", layer="rivs_CA_OR_hydroshed") 

# check crs
st_crs(rivers_sf)

```

## Polygon Shapefile Data

No different here, process is the same. But let's take a look at a package that might be helpful for folks working with state/county boundaries.

### Download State/County Data

A nice package for pulling county/state data is the `USAboundaries` package. Importantly, this package pulls these data in as `sf` features (`dataframes`), not as `rgdal` or `SpatialPolygonDataFrames` data.

Let's show this in two steps, the first is how to grab a `sf` feature for a given state or states.

```{r getStateData, echo=T, eval=T}

library(USAboundaries)

# Pick a State
state_names <- c("california")

# Download STATE data and add projection
CA<-us_states(resolution = "high", states = state_names) #%>%
  #st_transform(crs = 32610)

st_crs(CA)

```

That was easy...what about counties? We can use the same type of call, but let's add some `dplyr` and `purrr` functionality here to add the X and Y values for the centroid of each polygon (county) we download. In this case we use `map_dbl` because it will take a vector or values (the `geometry` col here), map a function over each row in that vector, and return a vector of values (the centroid points).

```{r getCountiesData, echo=T, eval=T}

library(purrr)

# Pick some CA counties
co_names <- c("Butte", "Placer", "El Dorado", "Nevada", "Yuba", "Sierra", "Plumas")

# get COUNTY data for a given state
counties_spec <- us_counties(resolution = "low", states=state_names) %>% # use list of state(s) here
  
  filter(name %in% co_names) %>% # filter to just the counties we want
  
  mutate(lon=map_dbl(geometry, ~st_centroid(.x)[[1]]), # add centroid values for labels
         lat=map_dbl(geometry, ~st_centroid(.x)[[2]])) # add centroid values for labels

```

## Make some Maps!

First let's show a few examples comparing how to plot with the `rgdal` rivers dataset vs. the `sf` version. 

### Base: `rgdal` vs. `sf`

```{r rgdal_base, echo=T, eval=T}

par(mfrow=c(1,2))
plot(rivers_sp, col="blue")
title("with sp")
plot(st_geometry(rivers_sf), col="darkblue")
title("with sf")

```

Pretty much the same deal. 

## Advanced Mapping

Okay, let's add in the county/state data, and figure out a few tricks to making our map a bit cleaner.

### Layering Maps: `plot` and `sf`

```{r plot_sf, eval=T, echo=T}
par(mfrow=c(1,1)) # reset to default

plot(st_geometry(counties_spec), col=adjustcolor("maroon", alpha=0.5), cex=1.5)
plot(st_geometry(rivers_sf), col=adjustcolor("blue", alpha=0.7), add=T)
text(counties_spec$lon, counties_spec$lat, labels = counties_spec$name)
graphics::title("Selected CA Counties")

```

Unforunately this map isn't very good. How can we improve it? Seems it would be nicer to crop the river layer to the counties of interest, or at least center things on that area. Also might be nice to have the labels on top and not obscured by the rivers.

```{r relayering}

# get range of lat/longs from counties for mapping
mapRange1 <- c(range(st_coordinates(counties_spec)[,1]),range(st_coordinates(counties_spec)[,2]))

plot(st_geometry(rivers_sf), col=adjustcolor("blue", alpha=0.7),
     xlim=mapRange1[c(1:2)], ylim = mapRange1[c(3:4)],axes=T)
plot(st_geometry(counties_spec), col=adjustcolor("maroon", alpha=0.5), cex=1.5, add=T)
text(counties_spec$lon, counties_spec$lat, labels = counties_spec$name,
     col=adjustcolor("white", alpha=0.8), font = 2)
graphics::title("Selected CA Counties")

```

That's a little better...but still not quite what we want. Lots of extra ink going toward stuff that we don't need (i.e., rivers outside of counties). What about cropping the river layer to our county layer?

### Cropping & Buffering with `sf`

Great thing is that `sf` has some really nice tools for this, just as any GIS program would.  Here let's use `st_intersection` to crop our river layer to only rivers in the counties of interest.

```{r crop_sf}

# crop rivers by counties
rivers_crop <- st_intersection(rivers_sf, counties_spec)

# plot
plot(st_geometry(rivers_crop), col="blue") # a quick plot check to make sure it worked

# buffer a single county?
county_buff <- st_buffer(counties_spec[counties_spec$name=="El Dorado",], dist = .05)
plot(st_geometry(counties_spec))
plot(st_geometry(county_buff), border="red", add=T)

```

Great, let's try our plot again.

```{r cropRelayer}

# try again, let's switch layer ordering
plot(st_geometry(counties_spec), col=adjustcolor("maroon", alpha=0.2), cex=1.5, axes=T)
plot(st_geometry(rivers_crop), col=adjustcolor("blue", alpha=0.7),add=T)
plot(st_geometry(CA), add=T, lwd=2)
text(counties_spec$lon, counties_spec$lat, labels = counties_spec$name,
     col="maroon", font = 2)
graphics::title("Selected CA Counties")
```

Ok! Not bad. What can we improve?

### `ggplot2` and `sf`

Mapping with ggplot2 brings some extra things we can fiddle with. Since these data are all data.frames (`sf` features), we can use the `geom_sf` function in the most recent version of `ggplot2`. I'm showing what happens by limiting the map range, vs. without the `coord_sf` call.

```{r ggplot_map1}

# not cropped
ggplot() + 
  geom_sf(data=CA, color = "gray30", lwd=2, fill=NA) +
  geom_sf(data=counties_spec, fill = NA, show.legend = F, color="gray50", lwd=0.4) +
  geom_label_repel(data=counties_spec, aes(x=lon, y=lat, label=name)) +
  theme_bw()

# with cropped range (to only our selected counties)
ggplot() + 
  geom_sf(data=CA, color = "gray30", lwd=2, fill=NA) +
  geom_sf(data=counties_spec, fill = NA, show.legend = F, color="gray50", lwd=0.4) +
  geom_sf(data=rivers_crop, col="blue", alpha=0.8, size=0.5)+
  geom_label_repel(data=counties_spec, aes(x=lon, y=lat, label=name)) +
  coord_sf(xlim = mapRange1[c(1:2)], ylim = mapRange1[c(3:4)]) +
  theme_bw(base_family = "Roboto Condensed") + # change to "sans" if this font not available
  labs(title="Selected CA Counties and Riverlines")

```

### Add ggmap

```{r}

library(ggmap)
location=c(-121,39.5) # set the center of the map
# set the background map up
map1 <- get_map(location=location, crop = F,
             color="bw",
             maptype="terrain",
             source="google",
             zoom=8)


ggmap(map1) +
  geom_sf(data=CA, color = "gray30", lwd=2, fill=NA, inherit.aes = F) +
  geom_sf(data=counties_spec, fill = NA, show.legend = F, color="gray50", lwd=0.4, inherit.aes = F) +
  geom_sf(data=rivers_crop, col="blue", alpha=0.8, size=0.5, inherit.aes = F)+
  geom_label_repel(data=counties_spec, aes(x=lon, y=lat, label=name)) +
  coord_sf(xlim = mapRange1[c(1:2)], ylim = mapRange1[c(3:4)]) +
  theme_bw(base_family = "Roboto Condensed") + # change to "sans" if this font not available
  labs(title="Selected CA Counties and Riverlines")
```

### Add an interactive *leaflet* map

We can add our `sf` data directly into a leaflet map using the `mapview` package.

```{r mapview}

library(mapview)

m1 <- mapview(rivers_crop)
m2 <- mapview(counties_spec, fill=NA)

# combine using the ggplot style "+"
m1 + m2
```


## Put it All Together

 - Can you crop to a single county and plot the rivers and county? 
 - How might you make an inset in your map? (hint...see [here](https://ryanpeek.github.io/2016-09-28-static_maps_in_R/))
 - What about buffering outside of the selected counties by 30 km?
 - Can you add some points? Try adding a point at **39.4 N and 121.0 W**.
