---
title: ""
---

<br>

# Working with `kml` Files

`kml` files are the unzipped versions of `kmz` files. You can use the `sf` package to read `kml`'s. You'll want to unzip with any sort of zip package you have on your computer (Keka, 7zip, etc.). Data used in this example live in the github repo [here](https://github.com/ryanpeek/mapping_in_R/blob/master/data/streamgages_06.kml).

```{r eval=T, echo=T}

suppressPackageStartupMessages({
  library(ggplot2)
  library(dplyr)
  #devtools::install_github("tidyverse/ggplot2", force=TRUE)
  library(sf);
  library(stringi)
  })

gages <- st_read("data/streamgages_06.kml") # reads kml but not kmz

```

### Cleaning/Removing `html` tags

Once we have the data in, we still need to clean up some of the `html` bits that we don't want. We can use some text/string functions to deal with this in R. `gsub` and `substr` are great functions for fixing/replacing text.

```{r eval=T, echo=T}

# pull out relevant info (remove html tags)
gages$name <- gsub(pattern="<.*?>", "", gages$Name)
gages$details <- gsub(pattern="<.*?>", "", gages$Description)
gages$site_no <- substr(gages$name, start = 1, stop = 8)
gages$site <- substr(gages$name, start = 10, stop = nchar(gages$name))
gages$date_s <- substr(gages$details, start=13, stop=22)
gages$date_e <- substr(gages$details, start=32, stop=nchar(gages$details)-1)

gages <- gages %>% dplyr::select(site_no:date_e)
gages$gageID <- paste0("T", gages$site_no)

# grep all sites with "AMERICAN", "YUBA" or "BEAR" in site name
gages_AMER <- filter(gages, grepl(pattern = "AMERICAN|YUBA|BEAR",site))

# simple map
ggplot() +
  labs(x="Longitude (WGS84)", y="Latitude",
       title="USGS Gages in CA") + 
  geom_sf(data=gages, col="blue", lwd=0.4, pch=21) +
  theme_bw(base_family = "Roboto Condensed") # change to sans if this font not local

```

### Add a `ggmap` background

Now we can add a different html background if we would like, or plot as a leaflet map with the `mapview` package.

```{r mapping, eval=T, echo=T}

# now with background
library(ggmap)

location<-c(mean(st_coordinates(gages_AMER)[,1]),
            mean(st_coordinates(gages_AMER)[,2]))

map1 <- get_map(location=location,crop = F,
                source = "stamen", maptype = "watercolor",
                zoom=8)
bg1 <-ggmap(map1)

map2 <- get_map(location=location, crop = F,
                source = "stamen", maptype = "toner",
                zoom=8)

bg2 <-ggmap(map2)

# water color version 
bg2 +
  scale_color_viridis_d() +
  labs(x="Longitude (WGS84)", y="Latitude",
       title="Selected USGS Gages with American | Yuba | Bear In Name") + 
  theme_bw(base_family = "Roboto Condensed") +
  geom_point(data=gages_AMER, 
             aes(x = st_coordinates(gages_AMER)[,1],
                 y = st_coordinates(gages_AMER)[,2]),
             fill="dodgerblue", alpha=0.8, pch=21, size=3)+
  coord_fixed(ratio=1.3)

```

